---
title: "Manipulating Network Data"
author: "Nic Fox"
date: "14/05/2021"
output: html_document
---

The code in this notebook is based on exercises in the book ["Statistical Analysis of Network Data with R" by Kolaczyk and Csardi](https://www.springer.com/gp/book/9781493909834). The authors provide all the [code for their book on Github](https://github.com/kolaczyk/sand).


# 2.2.1 Undirected & Directed Graphs

Terminology:

- **vertices** = nodes
- **edges** = links
- number of vertices (nodes) is called the '**order**' of the graph
- number of edges (links) is called the '**size**' of the graph
- 2 graphs that have the same structure and only differ in the naing of their vertices and edges are called **isomorphic**

Graphs can be **undirected** or **directed**.

A popular R package for working with network data is **igraph**.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("sand")  # created by the book authors & contains datasets used in the exercises
library(igraph)
```

## Create a graph

```{r create_first_graph}
# create the graph object
g <- graph_from_literal(1-2, 1-3, 2-3, 2-4, 3-5, 4-5, 4-6, 4-7, 5-6, 6-7)

# view the graph object details
g
```

## Visualise the graph



```{r visualise_first_graph}
# visualise the graph
plot(g)
```
```{r view_vertex_info}
V(g)
```

```{r view_edge_info}
E(g)
```

```{r print_all_graph_info}
print_all(g)
```

The **U** in the summary above indicates that the graph is **undirected**.

A **directed graph** is also called a **digraph** and edges in those types of graphs are called **directed edges** or **arcs**. Sometimes there may be 2 arcs between 2 vertices, each going in the opposite direction to the other, and these are called **mutual**.

```{r create_first_directed_graph}
# create the directed graph with 3 vertices and 4 edges (2 directed and one mutual)
dg <- graph_from_literal(1-+2, 1-+3, 2++3)

# view the information for the directed graph
dg
```

```{r visualise_first_directed_graph}
plot(dg)
```

```{r create_digraph_with_names}
dg_names <- graph_from_literal(Sam-+Mary, Sam-+Tom, Mary++Tom)
print_all(dg_names)
```

```{r visualise_digraph_with_names}
plot(dg_names)
```

```{r change_vertex_names}
V(dg_names)$name <- c("Sammy", "Mary Anne", "Tommy")
plot(dg_names)
```

# 2.2.2 Representations for Graphs

- adjacency list
- edge list
- adjacency matrix

## Adjacency list

Example: the list of nodes and list of directly associated nodes for each of those nodes, as seen in the output for `print_all()`

```{r view_adjacency_list_example}
print_all(g)
```


## Edge list

A 2-column list of vertex pairs that are joined by an edge.

```{r view_edge_list_example}
as_edgelist(g)
```


## Adjacency matrix

```{r view_adjacency_matrix_example}
as_adjacency_matrix(g)
```

## Useful functions

- `graph_from_adj_list()`
- `graph_from_edge_list()`
- `graph_from_adjacency_matrix()`
- `read_graph()` - for data stored in a file
- `write_graph()` - to save graph data to a file

# 2.2.3 Operations on Graphs

...